Binning Strategies and Related Loss for Large Data

### 2015-03-05 Paper.Rnw
Add section labels and start work on freq loss section

\section{Introduction}

@@

\section{Scatterplots for Large Data Sets}
In the case of modestly sized data, scatterplots are great tools for showing relationships in two dimensions. For large data, scatterplots suffer from over-plotting -- i.e. \ktm{a large number of observations drawn in such close proximity that their points overlap}, thereby masking relevant structure.
Figure~\ref{fig:scatter-alpha} shows an example taken from baseball statistics. The scatterplot shows 139 seasons (from the years of 1871 -- 2009) of  pitching statistics for every baseball pitcher as published in Sean Lahman's Baseball database (\url{http://www.seanlahman.com/baseball-archive/}).  The number of games played in a season is plotted versus number of strikeouts a pitcher threw over the course of a season. While the data set is only medium sized with \Sexpr{dim(d1.ext)[1]} observations, it already shows some of the break-down patterns scatterplots experience with large data.

@@

\section{Binning Algorithms}

@@

\subsection{Extension to Two Dimensional Binning}

@@

\section{Loss due to Binning}
While binning data has some notable advantages, it does result in loss of some information. Using a small number of bins removes much of the relevant information in the data, while an extremely large number of bins may not reduce the data size or complexity to a sufficient degree. 
Figure~\ref{binning} gives an overview of a data set and binned representations using different numbers of bins, demonstrating the loss of information with increasing bin size.

@@

\subsection{Spatial Loss}

@@

\subsection{Frequency Loss}

\ktm{If the shading scheme for rendering the counts in each bin is discretized there is a loss of frequency information. While counts can certainly be modeled as continuous on the HSV color space (reference needed) discrete color breaks may be preferred for a few reasons. %The average person is able to  On the continuous scale, a viewer will likely not be able to exactly match a tile shading to the corresponding count on the color key in the legend of the plot without an aid such as interactive mouseover effects. 
}

>>>>

\section{Introduction}
\label{Intro}

@@

\section{Scatterplots for Large Data Sets}
\label{Scatter}

In the case of modestly sized data, scatterplots are great tools for showing relationships in two dimensions. For large data, scatterplots suffer from over-plotting -- i.e. \ktm{a large number of observations drawn in such close proximity that their points overlap}, thereby masking relevant structure.
Figure~\ref{fig:scatter-alpha} shows an example taken from baseball statistics. The scatterplot shows 139 seasons (from the years of 1871 -- 2009) of  pitching statistics for every baseball pitcher as published in Sean Lahman's Baseball database (\url{http://www.seanlahman.com/baseball-archive/}).  The number of games played in a season is plotted versus number of strikeouts a pitcher threw over the course of a season. While the data set is only medium sized with \Sexpr{dim(d1.ext)[1]} observations, it already shows some of the break-down patterns scatterplots experience with large data.

@@

\section{Binning Algorithms}
\label{GenBinning}

@@

\subsection{Extension to Two Dimensional Binning}
\label{TwoDimBin}

@@

\section{Loss due to Binning}
\label{LossIntro}

While binning data has some notable advantages, it does result in loss of some information. Using a small number of bins removes much of the relevant information in the data, while an extremely large number of bins may not reduce the data size or complexity to a sufficient degree. 
Figure~\ref{binning} gives an overview of a data set and binned representations using different numbers of bins, demonstrating the loss of information with increasing bin size.

@@

\subsection{Spatial Loss}
\label{SpatialLoss}

@@

\subsection{Frequency Loss}
\label{FreqLoss}

\ktm{ If the shading scheme for rendering the counts in each bin is discretized there is a loss of frequency information. This is effectively a second stage of binning; wherein the bin counts, $c_\ell$, for bins $\ell \in \{1, \dots, L \}$ are placed into frequency bins  using any of the previously mentioned univariate binning algorithms in Section~\ref{GenBinning}. }

\ktm{ While counts can certainly be modeled as continuous on the HSV color space (reference needed) discrete color breaks may be preferred for a few reasons. The average person is able to distinguish  On the continuous scale, a viewer will likely not be able to exactly match a tile shading to the corresponding count on the color key in the legend of the plot without an aid such as interactive mouseover effects. 
}
